### 2025-10-08 YEAR2/AER210_LectureNotes/main.tex
oct 7

@@

@@

    Since $f(x,y)$ is seperable, we can write:
    $$
        \iint_R f(x,y) \, dA = \left( \int_0^{\frac{\pi}{2}} \sin{x} \, dx \right) \left( \int_0^{\frac{\pi}{2}} \cos{y} \, dy \right)
    $$
    Evaluating each integral separately gives 1 for both, so the final result is: $1 \times 1 = 1$.
\end{example}
\subsection{Double Integrals in General Regions}
\paragraph{Types of Regions} When the region $R$ is not rectangular, we can still compute the double integral by expressing the region in terms of inequalities. There are three common types of regions:
\begin{definition}[Type I Region]
    A region $R$ is called a Type I region if it can be described by the inequalities:
    $$
        a \le x \le b, \quad g_1(x) \le y \le g_2(x)
    $$
    where $g_1(x)$ and $g_2(x)$ are continuous functions on the interval $[a,b]$. Then, to evaluate the double integral over a Type I region for a continuous function $f(x,y)$, we set up the integral as follows:

@@

@@

\begin{definition}[Reigion Defined by Varying $r$ with $\theta$]
    Consider a region $R$ in the $xy$-plane that can be described in polar coordinates by the inequalities:
    $$
        \alpha \le \theta \le \beta, \quad g_1(\theta) \le r \le g_2(\theta)
    $$
    where $g_1(\theta)$ and $g_2(\theta)$ are continuous functions on the interval $[\alpha, \beta]$. Then, to evaluate the double integral over this region for a continuous function $f(x,y)$, we set up the integral as follows:
    \begin{equation}
        \iint_R f(x,y) \, dA = \int_\alpha^\beta \int_{g_1(\theta)}^{g_2(\theta)} f(r \cos{\theta}, r \sin{\theta}) \, r \, dr \, d\theta
    \end{equation}
\end{definition}

@@

\subsection{Applications of Double Integrals}
\begin{shaded}
    \subsubsection*{Review: Moment of Inertia using Single Integral}

@@

\subsection{Surface Area}
\begin{theorem}[Surface Area]
    Given $z=f(x,y)$, where $f$ is a differentiable function over the region $R$ in the $xy$-plane, the surface area $S$ of the surface above the region $R$ is given by:
    \begin{equation}
        S = \iint_R \sqrt{1 + \left( \frac{\partial f}{\partial x} \right)^2 + \left( \frac{\partial f}{\partial y} \right)^2} \, dA
    \end{equation}
\end{theorem}
\begin{proof}
    Consider a small rectangle $\Delta A_i$ in the $xy$-plane with dimensions $\Delta x_i$ and $\Delta y_i$. When we project this rectangle onto the surface $z = f(x,y)$, it becomes a parallelogram $T_i$ tangent to the surface. Using the partial derivatives of the surface, we can deduce the two vectors that define the parallelogram:
    $$
        \vec{u} = \left( \Delta x_i, 0, f_x(x_i^*, y_i^*) \Delta x_i \right), \quad \vec{v} = \left( 0, \Delta y_i, f_y(x_i^*, y_i^*) \Delta y_i \right)
    $$
    The area of this parallelogram is given by the magnitude of the cross product of these two vectors:
    \begin{align*}
        \Delta T_i = ||\vec{u} \times \vec{v}|| &= \left|\begin{vmatrix}
        \hat{i} & \hat{j} & \hat{k} \\
        \Delta x_i & 0 & f_x(x_i^*, y_i^*) \Delta x_i \\
        0 & \Delta y_i & f_y(x_i^*, y_i^*) \Delta y_i
        \end{vmatrix}\right| \\
        &= \sqrt{( -f_x(x_i^*, y_i^*) \Delta x_i \Delta y_i )^2 + ( -f_y(x_i^*, y_i^*) \Delta x_i \Delta y_i )^2 + (\Delta x_i \Delta y_i)^2} \\
        &= \sqrt{1 + (f_x(x_i^*, y_i^*))^2 + (f_y(x_i^*, y_i^*))^2} \, \Delta x_i \Delta y_i
    \end{align*}
    Therefore, the surface area can be approximated as:
    $$
        S \approx S_n = \sum_{i=1}^n \Delta T_i = \sum_{i=1}^n \sqrt{1 + (f_x(x_i^*, y_i^*))^2 + (f_y(x_i^*, y_i^*))^2} \, \Delta A_i
    $$
    Taking the limit as the maximum diameter of the subrectangles approaches zero, we obtain the exact surface area:
    $$
        S = \lim_{|P| \to 0} S_n =\iint_R \sqrt{1 + \left( \frac{\partial f}{\partial x} \right)^2 + \left( \frac{\partial f}{\partial y} \right)^2} \, dA
    $$
    which completes our derivation.
\end{proof}

@@

\subsection{Triple Integrals in Rectangular Coordinates}
The idea of a triple integral (in fact n-tuple integral) could be extend from the idea of double integral, similar to 
Definition \ref{def:double_integral}. We can define a triple integral as follows:
\begin{definition}
    Consider a function $f(x,y,z)$ that is continuous a 
    3-D reigion with volumne $V$. We can partition the reigion into $n$ subreigons with volume $\Delta V_i$. Then, we can define the triple integral of $f$ over the reigion $V$ as follows:
    \begin{equation}
        \iiint_V f(x,y,z) \, dV = \lim_{|P| \to 0} \sum_{i=1}^n f(x_i^*, y_i^*, z_i^*) \Delta V_i
    \end{equation}
    where $(x_i^*, y_i^*, z_i^*)$ is a point in the $i$-th subregion, and $|P|$ is the maximum diameter of the subregions. This limit exists and equals to the triple integral if $f$ is continuous on $V$, since the following squeeze theorem holds, similar to the double integral case:
    $$       
     \sum_{i=1}^n m_i \Delta V_i \le \sum_{i=1}^n f(x_i^*, y_i^*, z_i^*) \Delta V_i \le \sum_{i=1}^n M_i \Delta V_i
    $$
    where $m$ and $M$ are the minimum and maximum values of $f$ on $V$, respectively.

@@

    Then, we can express the base region $R$ in polar coordinates:
    $$        
    x^2 + y^2 = r^2 \implies 0 \le r \le a, \quad 0 \le \theta \le 2\pi
    $$
    So we can set up:
    \begin{align*}
        I_z &= \rho_0 h \int_0^{2\pi} \int_0^a r^2 \cdot r \, dr \, d\theta \\
        &= \rho_0 h \int_0^{2\pi} \left[ \frac{r^4}{4} \right]_{r=0}^{r=a} \, d\theta = \frac{\pi \rho_0 h a^4}{2}
    \end{align*}
    This also leads to the well-known formula for the moment of inertia of a solid cylinder about its central axis:
    $$
        I_z = \frac{1}{2} M a^2
    $$
    where $M = \pi a^2 h \rho_0$ is the mass of the cylinder. 
\end{example}
\subsection{Triple Integrals in Cylindrical Coordinates}
\begin{definition}[Cylindrical Coordinates]
    The cylindrical coordinates of a point $P$ in 3-D space are given by the ordered triple $(r, \theta, z)$, where:
    \begin{itemize}
        \item $r$ is the distance from the $z$-axis to the projection of $P$ onto the $xy$-plane,
        \item $\theta$ is the angle between the positive $x$-axis and the line segment from the origin to the projection of $P$ onto the $xy$-plane,
        \item $z$ is the same as in rectangular coordinates, representing the height of point $P$ above the $xy$-plane.
    \end{itemize}
    The relationships between cylindrical coordinates and rectangular coordinates are given by:
    \begin{equation}
        x = r \cos \theta, \quad y = r \sin \theta, \quad z = z
    \end{equation}
    and conversely:
    \begin{equation}
        r = \sqrt{x^2 + y^2}, \quad \theta = \tan^{-1}\left(\frac{y}{x}\right), \quad z = z
    \end{equation}

@@

\end{example}
\subsection{Triple Integrals in Spherical Coordinates}
\begin{definition}[Spherical Coordinates]
    The spherical coordinates of a point $P$ in 3-D space are given by the ordered triple $(\rho, \phi, \theta)$, where:
    \begin{itemize}
        \item $\rho$ is the distance from the origin to the point $P$,
        \item $\phi$ is the angle between the positive $z$-axis and the line segment from the origin to point $P$ (also known as the polar angle or colatitude),
        \item $\theta$ is the angle between the positive $x$-axis and the projection of the line segment from the origin to point $P$ onto the $xy$-plane (also known as the azimuthal angle).
    \end{itemize}
    The relationships between spherical coordinates and rectangular coordinates are given by:
    \begin{equation}
        x = \rho \sin \phi \cos \theta, \quad y = \rho \sin \phi \sin \theta, \quad z = \rho \cos \phi
    \end{equation}
    and conversely:
    \begin{equation}
        \rho = \sqrt{x^2 + y^2 + z^2}, \quad \phi = \cos^{-1}\left(\frac{z}{\sqrt{x^2 + y^2 + z^2}}\right), \quad \theta = \tan^{-1}\left(\frac{y}{x}\right)
    \end{equation}

@@

\end{example}
\subsection{Change of Variables in Multiple Integrals} \label{sec:change_of_variables}
\begin{shaded}
    \subsubsection{Change of Variables in Single Integrals}
    Consider the following integral:
    $$
        \int_1^3 2x\sqrt{x^2 +1} \, dx
    $$
    We can use substitution to solve this integral. Let $u = x^2 + 1$, then $du = 2x \, dx$. Then, we carefully plug in to the integral, and notice how this could extend to multiple integrals:
    \begin{align*}
        \int_1^3 2x\sqrt{x^2 +1} \, dx &= \int_{u=2}^{u=10} 2 \sqrt{u} \sqrt{u-1} \cdot \frac{du}{2\sqrt{u-1}} = \int_{u=2}^{u=10} \sqrt{u} \, du \\
        &= \left[ \frac{2}{3} u^{\frac{3}{2}} \right]_{u=2}^{u=10} = \frac{2}{3} (10^{\frac{3}{2}} - 2^{\frac{3}{2}})
    \end{align*}
    The key step is to express $dx$ in terms of $du$:
    $$        
    dx = \frac{du}{2x}
    $$
    Consider how the area under a curve shifts when we change the variable from $x$ to $u$, we can visualize this as follows:
    % two minipages side by side one with the graph of y = f(x) from 0 to 3 and the other with the graph of y = f(u) from 2 to 10
    \end{shaded}
    \newpage

@@

\section{Vector Calculus}
\section{Line Integrals}
\begin{definition}[Line Integral of a Scalar Field]
    Consider a scalar field $f(x,y) $ and a smooth curve $C$. Then we define the line integral of $f$ along $C$ as the following Riemann sum:
    $$
        \int_C f(x,y) \, ds = \lim_{||P|| \to 0} \sum_{i=1}^n f(x_i^*, y_i^*) \Delta s_i
    $$
    where $P$ is a partition of $C$ into $n$ subintervals, $||P||$ is the norm of the partition, $(x_i^*, y_i^*)$ is a sample point in the $i$-th subinterval, and $\Delta s_i$ is the length of the $i$-th subinterval.
\end{definition}

@@

\begin{theorem}[Directionality of Scalar Line Integrals] The line integral of a scalar field is independent of the direction of the curve. That is, if we reverse the direction of the curve, the value of the line integral remains the same:
\begin{equation}
    \int_C f(x,y) \, ds = \int_{-C} f(x,y) \, ds
\end{equation}
\end{theorem}
\begin{proof}
    Let $C$ be a smooth curve parametrized by $\vec{r(t)} = \langle g(t), h(t) \rangle$, $a \leq t \leq b$. Then, the line integral of $f$ along $C$ is given by:
    \begin{equation}
        \int_C f(x,y) \, ds = \int_a^b f(g(t), h(t)) \, ||\vec{r'(t)}|| \, dt
    \end{equation}
    Now, consider the curve $-C$ obtained by reversing the direction of $C$. We can parametrize $-C$ by $\vec{r(-t)} = \langle g(-t), h(-t) \rangle$, $-b \leq t \leq -a$. The line integral of $f$ along $-C$ is given by:
    \begin{align*}
        \int_{-C} f(x,y) \, ds &= \int_{-b}^{-a} f(g(-t), h(-t)) \, ||\vec{r'(-t)}|| \, dt \\
        &= \int_a^b f(g(u), h(u)) \, ||\vec{r'(u)}|| \, du \quad (u = -t, du = -dt) \\
        &= \int_a^b f(g(t), h(t)) \, ||\vec{r'(t)}|| \, dt
    \end{align*}
    Thus, we have shown that:
    $$
        \int_C f(x,y) \, ds = \int_{-C} f(x,y) \, ds
    $$
\end{proof}
\chapter{Fluid Mechanics}
\end{document}

>>>>

\begin{keybox}
	\textbf{Key Equations:}
\begin{itemize}
    \item Multivariable Taylor (about $(x_0,y_0)$): $\displaystyle f(x,y) = f_0 + \sum_i f_i\Delta x_i + \tfrac12\sum_{i,j} f_{ij}\Delta x_i\Delta x_j + \cdots$
    \item 2D form: $\displaystyle f(x_0{+}\Delta x,y_0{+}\Delta y) = f_0 + f_x\Delta x + f_y\Delta y + \tfrac12(f_{xx}\Delta x^2 + 2f_{xy}\Delta x\Delta y + f_{yy}\Delta y^2)+\cdots$
\end{itemize}
	\textbf{Tips:}
\begin{itemize}
    \item Use a line path $(x,y)=(x_0,y_0)+t(\Delta x,\Delta y)$ to reduce to 1D.
    \item Hessian controls the quadratic term and local shape (min/max/saddle with tests later).
\end{itemize}
\end{keybox}

@@

\begin{keybox}
	\textbf{Key Equations:}
\begin{itemize}
    \item Rectangle $R=[a,b]\times[c,d]$: $\displaystyle \iint_R f\,dA = \int_a^b\!\int_c^d f(x,y)\,dy\,dx = \int_c^d\!\int_a^b f(x,y)\,dx\,dy$
    \item Separable $f(x,y)=g(x)h(y)$: product of one-dimensional integrals.
\end{itemize}
	\textbf{Tips:}
\begin{itemize}
    \item Swap order freely on rectangles; use Fubini/Tonelli conditions.
    \item Check integrability and symmetry to simplify.
\end{itemize}
\end{keybox}

@@

    Since $f(x,y)$ is seperable, we can write:
    $$
        \iint_R f(x,y) \, dA = \left( \int_0^{\frac{\pi}{2}} \sin{x} \, dx \right) \left( \int_0^{\frac{\pi}{2}} \cos{y} \, dy \right)
    $$
    Evaluating each integral separately gives 1 for both, so the final result is: $1 \times 1 = 1$.
\end{example}
\subsection{Double Integrals in General Regions}

\begin{keybox}
	\textbf{Key Ideas:}
\begin{itemize}
    \item Type I: $a\le x\le b$, $g_1(x)\le y\le g_2(x)$, integrate $dy$ then $dx$.
    \item Type II: $c\le y\le d$, $h_1(y)\le x\le h_2(y)$, integrate $dx$ then $dy$.
\end{itemize}
	\textbf{Tips/Pitfalls:}
\begin{itemize}
    \item Draw the region; switching order often simplifies bounds.
    \item Split region if boundaries change formula across sub-intervals.
\end{itemize}
\end{keybox}
\paragraph{Types of Regions} When the region $R$ is not rectangular, we can still compute the double integral by expressing the region in terms of inequalities. There are three common types of regions:
\begin{definition}[Type I Region]
    A region $R$ is called a Type I region if it can be described by the inequalities:
    $$
        a \le x \le b, \quad g_1(x) \le y \le g_2(x)
    $$
    where $g_1(x)$ and $g_2(x)$ are continuous functions on the interval $[a,b]$. Then, to evaluate the double integral over a Type I region for a continuous function $f(x,y)$, we set up the integral as follows:

@@

\begin{keybox}
	\textbf{Key Equations/Concepts:}
\begin{itemize}
    \item Piecewise descriptions: break $R$ into Type I/II parts.
    \item Use symmetry and monotonic boundaries to choose efficient order.
\end{itemize}
	\textbf{Tip:} Convert to polar if circular arcs or $x^2{+}y^2$ appear.
\end{keybox}

@@

\begin{definition}[Reigion Defined by Varying $r$ with $\theta$]
    Consider a region $R$ in the $xy$-plane that can be described in polar coordinates by the inequalities:
    \begin{equation}
        \alpha \le \theta \le \beta, \quad g_1(\theta) \le r \le g_2(\theta)
    \end{equation}
    where $g_1(\theta)$ and $g_2(\theta)$ are continuous functions on the interval $[\alpha, \beta]$. Then, to evaluate the double integral over this region for a continuous function $f(x,y)$, we set up the integral as follows:
    \begin{equation}
        \iint_R f(x,y) \, dA = \int_\alpha^\beta \int_{g_1(\theta)}^{g_2(\theta)} f(r \cos{\theta}, r \sin{\theta}) \, r \, dr \, d\theta
    \end{equation}
\end{definition}

@@

\subsection{Applications of Double Integrals}

\begin{keybox}
	\textbf{Key Formulas:}
\begin{itemize}
    \item Mass: $M=\iint_R \rho\,dA$; Center of mass: $\bar x=\frac{1}{M}\iint_R x\rho\,dA$, $\bar y=\frac{1}{M}\iint_R y\rho\,dA$.
    \item Moments of inertia: $I_x=\iint_R y^2\rho\,dA$, $I_y=\iint_R x^2\rho\,dA$, $I_O=I_x+I_y$ (planar).
\end{itemize}
	\textbf{Tips:}
\begin{itemize}
    \item Use symmetry to zero cross-terms and simplify integrals.
    \item Choose coordinates aligned with geometry.
\end{itemize}
\end{keybox}
\begin{shaded}
    \subsubsection*{Review: Moment of Inertia using Single Integral}

@@

\subsection{Surface Area}

\begin{keybox}
	\textbf{Key Formula:}
\begin{itemize}
    \item Graph $z=f(x,y)$ over $R$: $\displaystyle S=\iint_R \sqrt{1+f_x^2+f_y^2}\,dA$.
\end{itemize}
	\textbf{Tips:}
\begin{itemize}
    \item For parametric $\mathbf r(u,v)$ use $\iint\!\lVert \mathbf r_u\times\mathbf r_v\rVert\,du\,dv$.
    \item Approximate with linearization when small slopes.
\end{itemize}
\end{keybox}
\begin{theorem}[Surface Area]
    Given $z=f(x,y)$, where $f$ is a differentiable function over the region $R$ in the $xy$-plane, the surface area $S$ of the surface above the region $R$ is given by:
    \begin{equation}
        S = \iint_R \sqrt{1 + \left( \frac{\partial f}{\partial x} \right)^2 + \left( \frac{\partial f}{\partial y} \right)^2} \, dA
    \end{equation}
\end{theorem}
\begin{proof}
    Consider a small rectangle $\Delta A_i$ in the $xy$-plane with dimensions $\Delta x_i$ and $\Delta y_i$. When we project this rectangle onto the surface $z = f(x,y)$, it becomes a parallelogram $T_i$ tangent to the surface. Using the partial derivatives of the surface, we can deduce the two vectors that define the parallelogram:
    $$
        \vec{u} = \left( \Delta x_i, 0, f_x(x_i^*, y_i^*) \Delta x_i \right), \quad \vec{v} = \left( 0, \Delta y_i, f_y(x_i^*, y_i^*) \Delta y_i \right)
    $$
    The area of this parallelogram is given by the magnitude of the cross product of these two vectors:
    \begin{align*}
        \Delta T_i = ||\vec{u} \times \vec{v}|| &= \left|\begin{vmatrix}
        \hat{i} & \hat{j} & \hat{k} \\
        \Delta x_i & 0 & f_x(x_i^*, y_i^*) \Delta x_i \\
        0 & \Delta y_i & f_y(x_i^*, y_i^*) \Delta y_i
        \end{vmatrix}\right| \\
        &= \sqrt{( -f_x(x_i^*, y_i^*) \Delta x_i \Delta y_i )^2 + ( -f_y(x_i^*, y_i^*) \Delta x_i \Delta y_i )^2 + (\Delta x_i \Delta y_i)^2} \\
        &= \sqrt{1 + (f_x(x_i^*, y_i^*))^2 + (f_y(x_i^*, y_i^*))^2} \, \Delta x_i \Delta y_i
    \end{align*}
    Therefore, the surface area can be approximated as:
    $$
        S \approx S_n = \sum_{i=1}^n \Delta T_i = \sum_{i=1}^n \sqrt{1 + (f_x(x_i^*, y_i^*))^2 + (f_y(x_i^*, y_i^*))^2} \, \Delta A_i
    $$
    Taking the limit as the maximum diameter of the subrectangles approaches zero, we obtain the exact surface area:
    $$
        S = \lim_{|P| \to 0} S_n =\iint_R \sqrt{1 + \left( \frac{\partial f}{\partial x} \right)^2 + \left( \frac{\partial f}{\partial y} \right)^2} \, dA
    $$
    which completes our derivation.
\end{proof}

@@

\subsection{Triple Integrals in Rectangular Coordinates}

\begin{keybox}
	\textbf{Key Equations:}
\begin{itemize}
    \item $\displaystyle \iiint_E f\,dV = \int\!\int\!\int f(x,y,z)\,dz\,dy\,dx$ over box or layered bounds.
    \item Mass/CM/ Moments extend by adding $z$ and $dV$.
\end{itemize}
	\textbf{Tips:}
\begin{itemize}
    \item Choose order by inspecting projection and boundary dependence.
    \item Slice by planes to keep inner bounds simple.
\end{itemize}
\end{keybox}
The idea of a triple integral (in fact n-tuple integral) could be extend from the idea of double integral, similar to 
Definition \ref{def:double_integral}. We can define a triple integral as follows:
\begin{definition}
    Consider a function $f(x,y,z)$ that is continuous a 
    3-D reigion with volumne $V$. We can partition the reigion into $n$ subreigons with volume $\Delta V_i$. Then, we can define the triple integral of $f$ over the reigion $V$ as follows:
    \begin{equation}
        \iiint_V f(x,y,z) \, dV = \lim_{|P| \to 0} \sum_{i=1}^n f(x_i^*, y_i^*, z_i^*) \Delta V_i
    \end{equation}
    where $(x_i^*, y_i^*, z_i^*)$ is a point in the $i$-th subregion, and $|P|$ is the maximum diameter of the subregions. This limit exists and equals to the triple integral if $f$ is continuous on $V$, since the following squeeze theorem holds, similar to the double integral case:
    $$       
     \sum_{i=1}^n m_i \Delta V_i \le \sum_{i=1}^n f(x_i^*, y_i^*, z_i^*) \Delta V_i \le \sum_{i=1}^n M_i \Delta V_i
    $$
    where $m$ and $M$ are the minimum and maximum values of $f$ on $V$, respectively.

@@

    Then, we can express the base region $R$ in polar coordinates:
    $$        
    x^2 + y^2 = r^2 \implies 0 \le r \le a, \quad 0 \le \theta \le 2\pi
    $$
    So we can set up:
    \begin{align*}
        I_z &= \rho_0 h \int_0^{2\pi} \int_0^a r^2 \cdot r \, dr \, d\theta \\
        &= \rho_0 h \int_0^{2\pi} \left[ \frac{r^4}{4} \right]_{r=0}^{r=a} \, d\theta = \frac{\pi \rho_0 h a^4}{2}
    \end{align*}
    This also leads to the well-known formula for the moment of inertia of a solid cylinder about its central axis:
    $$
        I_z = \frac{1}{2} M a^2
    $$
    where $M = \pi a^2 h \rho_0$ is the mass of the cylinder. 
\end{example}
\subsection{Triple Integrals in Cylindrical Coordinates}

\begin{keybox}
	\textbf{Key Relations:}
\begin{itemize}
    \item $(r,\theta,z)$ with $x=r\cos\theta$, $y=r\sin\theta$, $dV=r\,dr\,d\theta\,dz$.
    \item Use when symmetry around $z$-axis or $x^2{+}y^2$ appears.
\end{itemize}
	\textbf{Tip:} Don’t forget the Jacobian factor $r$.
\end{keybox}
\begin{definition}[Cylindrical Coordinates]
    The cylindrical coordinates of a point $P$ in 3-D space are given by the ordered triple $(r, \theta, z)$, where:
    \begin{itemize}
        \item $r$ is the distance from the $z$-axis to the projection of $P$ onto the $xy$-plane,
        \item $\theta$ is the angle between the positive $x$-axis and the line segment from the origin to the projection of $P$ onto the $xy$-plane,
        \item $z$ is the same as in rectangular coordinates, representing the height of point $P$ above the $xy$-plane.
    \end{itemize}
    The relationships between cylindrical coordinates and rectangular coordinates are given by:
    \begin{equation}
        x = r \cos \theta, \quad y = r \sin \theta, \quad z = z
    \end{equation}
    and conversely:
    \begin{equation}
        r = \sqrt{x^2 + y^2}, \quad \theta = \tan^{-1}\left(\frac{y}{x}\right), \quad z = z
    \end{equation}

@@

\end{example}
\subsection{Triple Integrals in Spherical Coordinates}

\begin{keybox}
	\textbf{Key Relations:}
\begin{itemize}
    \item $(\rho,\phi,\theta)$ with $x=\rho\sin\phi\cos\theta$, $y=\rho\sin\phi\sin\theta$, $z=\rho\cos\phi$.
    \item $dV=\rho^2\sin\phi\,d\rho\,d\phi\,d\theta$.
\end{itemize}
	\textbf{Tips:}
\begin{itemize}
    \item Spheres/cones/spherical caps are natural in these coordinates.
    \item Bound order: typically $\rho$ (inner), then $\phi$, then $\theta$.
\end{itemize}
\end{keybox}
\begin{definition}[Spherical Coordinates]
    The spherical coordinates of a point $P$ in 3-D space are given by the ordered triple $(\rho, \phi, \theta)$, where:
    \begin{itemize}
        \item $\rho$ is the distance from the origin to the point $P$,
        \item $\phi$ is the angle between the positive $z$-axis and the line segment from the origin to point $P$ (also known as the polar angle or colatitude),
        \item $\theta$ is the angle between the positive $x$-axis and the projection of the line segment from the origin to point $P$ onto the $xy$-plane (also known as the azimuthal angle).
    \end{itemize}
    The relationships between spherical coordinates and rectangular coordinates are given by:
    \begin{equation}
        x = \rho \sin \phi \cos \theta, \quad y = \rho \sin \phi \sin \theta, \quad z = \rho \cos \phi
    \end{equation}
    and conversely:
    \begin{equation}
        \rho = \sqrt{x^2 + y^2 + z^2}, \quad \phi = \cos^{-1}\left(\frac{z}{\sqrt{x^2 + y^2 + z^2}}\right), \quad \theta = \tan^{-1}\left(\frac{y}{x}\right)
    \end{equation}

@@

\end{example}
\subsection{Change of Variables in Multiple Integrals} \label{sec:change_of_variables}

\begin{keybox}
	\textbf{Key Equations:}
\begin{itemize}
    \item $\displaystyle \iint_R f(x,y)\,dA = \iint_{R'} f(x(u,v),y(u,v))\,\left|\det J\right|\,du\,dv$.
    \item $J=\partial(x,y)/\partial(u,v)$; similarly for 3D.
\end{itemize}
	\textbf{Tips/Pitfalls:}
\begin{itemize}
    \item Map the region carefully; sketch both $R$ and $R'$.
    \item Include the Jacobian; watch sign via absolute value.
\end{itemize}
\end{keybox}
\begin{shaded}
    \subsubsection{Change of Variables in Single Integrals}
    Consider the following integral:
    $$
        \int_1^3 2x\sqrt{x^2 +1} \, dx
    $$
    We can use substitution to solve this integral. Let $u = x^2 + 1$, then $du = 2x \, dx$. Then, we carefully plug in to the integral, and notice how this could extend to multiple integrals:
    \begin{align*}
        \int_1^3 2x\sqrt{x^2 +1} \, dx &= \int_{u=2}^{u=10} 2 \sqrt{u} \sqrt{u-1} \cdot \frac{du}{2\sqrt{u-1}} = \int_{u=2}^{u=10} \sqrt{u} \, du \\
        &= \left[ \frac{2}{3} u^{\frac{3}{2}} \right]_{u=2}^{u=10} = \frac{2}{3} (10^{\frac{3}{2}} - 2^{\frac{3}{2}})
    \end{align*}
    The key step is to express $dx$ in terms of $du$:
    $$        
    dx = \frac{du}{2x}
    $$
    Consider how the area under a curve shifts when we change the variable from $x$ to $u$, we can visualize this as follows:
    % two minipages side by side one with the graph of y = f(x) from 0 to 3 and the other with the graph of y = f(u) from 2 to 10
    \end{shaded}
    \newpage

@@

\section{Vector Calculus}
\section{Line Integrals}

\subsection{Line Integrals and Vector Fields}

\begin{keybox}
	\textbf{Key Equations:}
\begin{itemize}
    \item Scalar field: $\displaystyle \int_C f\,ds = \int_a^b f(\mathbf r(t))\,\lVert \mathbf r'(t)\rVert\,dt$.
    \item Vector field work: $\displaystyle \int_C \mathbf F\cdot d\mathbf r = \int_a^b \mathbf F(\mathbf r(t))\cdot\mathbf r'(t)\,dt$.
\end{itemize}
	\textbf{Tips:}
\begin{itemize}
    \item Orientation matters; reverse curve flips sign of vector line integral.
    \item For conservative fields, path independence and potential functions apply.
\end{itemize}
\end{keybox}
\begin{definition}[Line Integral of a Scalar Field]
    Consider a scalar field $f(x,y) $ and a smooth curve $C$. Then we define the line integral of $f$ along $C$ as the following Riemann sum:
    $$
        \int_C f(x,y) \, ds = \lim_{||P|| \to 0} \sum_{i=1}^n f(x_i^*, y_i^*) \Delta s_i
    $$
    where $P$ is a partition of $C$ into $n$ subintervals, $||P||$ is the norm of the partition, $(x_i^*, y_i^*)$ is a sample point in the $i$-th subinterval, and $\Delta s_i$ is the length of the $i$-th subinterval.
\end{definition}

@@

\begin{theorem}[Directionality of Scalar Line Integrals] The line integral of a scalar field is independent of the direction of the curve. That is, if we reverse the direction of the curve, the value of the line integral remains the same:
\begin{equation}
    \int_C f(x,y) \, ds = \int_{-C} f(x,y) \, ds
\end{equation}
\end{theorem}
\begin{proof}
    Let $C$ be a smooth curve parametrized by $\vec{r(t)} = \langle g(t), h(t) \rangle$, $a \leq t \leq b$. Then, the line integral of $f$ along $C$ is given by:
    \begin{equation}
        \int_C f(x,y) \, ds = \int_a^b f(g(t), h(t)) \, ||\vec{r'(t)}|| \, dt
    \end{equation}
    Now, consider the curve $-C$ obtained by reversing the direction of $C$. We can parametrize $-C$ by $\vec{r(-t)} = \langle g(-t), h(-t) \rangle$, $-b \leq t \leq -a$. The line integral of $f$ along $-C$ is given by:
    \begin{align*}
        \int_{-C} f(x,y) \, ds &= \int_{-b}^{-a} f(g(-t), h(-t)) \, ||\vec{r'(-t)}|| \, dt \\
        &= \int_a^b f(g(u), h(u)) \, ||\vec{r'(u)}|| \, du \quad (u = -t, du = -dt) \\
        &= \int_a^b f(g(t), h(t)) \, ||\vec{r'(t)}|| \, dt
    \end{align*}
    Thus, we have shown that:
    $$
        \int_C f(x,y) \, ds = \int_{-C} f(x,y) \, ds
    $$
\end{proof}

\begin{definition}[Line Integral of a Vector Field]
    Consider a vector field $\vec{F}(x,y)$ and a smooth curve $C$. Then we define the line integral of $\vec{F}$ along $C$ as the following Riemann sum:
    \begin{equation}
        \int_C \vec{F} \cdot d\vec{r} = \lim_{||P|| \to 0} \sum_{i=1}^n \vec{F}(x_i^*, y_i^*) \cdot \Delta \vec{r_i}
    \end{equation}
    where $P$ is a partition of $C$ into $n$ subintervals and $||P||$ is the norm of the partition, $(x_i^*, y_i^*)$ is a sample point in the $i$-th subinterval, and $\Delta \vec{r_i}$ is the vector from the start to the end of the $i$-th subinterval. We define $d \vec{r}$ as the infinitesimal vector along the curve:
    \begin{equation}
        d\vec{r} = \frac{d\vec{r}}{dt} dt = \vec{T} \, ds
    \end{equation}
    where $\vec{T}$ is the unit tangent vector to the curve and $ds$ is the infinitesimal arc length along the curve.

    To evaluate the line integral, we can use the parametrization of the curve $C$:
    \begin{equation}
        \int_C F \cdot d\vec{r} = \int_a^b \vec{F}(r(t)) \cdot \vec{r'(t)} \, dt
    \end{equation}
    where $\vec{r(t)} = \langle g(t), h(t) \rangle$, $a \leq t \leq b$ is the parametrization of the curve $C$.
\end{definition}

\begin{example}
    Consider the line integral of $\vec{F}(x, y, z) = yz \, \hat{i} + xz \, \hat{j} + xy \, \hat{k}$ along the curve $C$ parametrized by $ \vec{r}(t) = t^2 \, \hat{i} + t \, \hat{j} + t^3 \, \hat{k}, \quad 0 \leq t \leq 2 $. We have that:
    \begin{align*}
        \int_C \vec{F} \cdot d\vec{r} &= \int_0^2 \vec{F}(r(t)) \cdot \vec{r'(t)} \, dt \\
        &= \int_0^2 \langle t \cdot t^3, t^2 \cdot t^3, t^2 \cdot t \rangle \cdot \langle 2t, 1, 3t^2 \rangle \, dt \\
        &= \int_0^2 \langle t^4, t^5, t^3 \rangle \cdot \langle 2t, 1, 3t^2 \rangle \, dt \\
        &= \int_0^2 (2t^5 + t^5 + 3t^5) \, dt = \int_0^2 6t^5 \, dt = \left[ t^6 \right]_0^2 = 64
    \end{align*}
\end{example}

\subsection{Fundamental Theorem of Line Integrals}

\begin{keybox}
	\textbf{Key Statement:}
\begin{itemize}
    \item If $\mathbf F=\nabla \phi$ and $C$ goes from $A$ to $B$, then $\displaystyle \int_C \mathbf F\cdot d\mathbf r = \phi(B)-\phi(A)$.
\end{itemize}
	\textbf{Consequences:}
\begin{itemize}
    \item Path independence; integral on closed loops is $0$.
    \item Test for conservative in simply connected regions: $\nabla\times\mathbf F=\mathbf 0$.
\end{itemize}
\end{keybox}
\begin{shaded}
Recall, fundametnal theorem of calculus:
\begin{theorem}[Fundamental Theorem of Calculus]
    If $f$ is continuous on $[a,b]$ and $F$ is an antiderivative of $f$ on $[a,b]$, then:
    \begin{equation}
        \int_a^b f(x) \, dx = F(b) - F(a)
    \end{equation}
\end{theorem}
\end{shaded}

\begin{definition}[Conservative Vector Field]
    A vector field $\vec{F}$ is said to be conservative if there exists a scalar potential function $\phi$ such that:
    \begin{equation}
        \vec{F} = \nabla \phi
    \end{equation}
    where $\nabla \phi$ is the gradient of $\phi$.

\begin{theorem}[Test for Conservative Vector Field]
    A vector field $\vec{F} = P \, \hat{i} + Q \, \hat{j} + R \, \hat{k}$ is conservative if:
    \begin{equation}
        \frac{\partial P}{\partial y} = \frac{\partial Q}{\partial x}, \quad \frac{\partial P}{\partial z} = \frac{\partial R}{\partial x}, \quad \frac{\partial Q}{\partial z} = \frac{\partial R}{\partial y}
    \end{equation}
    in a simply connected domain.
\end{theorem}

\begin{theorem}[Conservative Closed Integral]
    If \(\vec F\) is a conservative vector field on an open, simply connected region \(D\), then for every closed curve \(C\) in \(D\),
    \[
        \oint_C \vec F \cdot d\vec r = 0
    \]
\end{theorem}
